\documentclass{pxcsdoc}
\usepackage{praxfuzz}
\usepackage{parskip}
\usepackage{multicol}
\usepackage{enumerate}
\usepackage{proof}


\newcommand{\true}{\mathrel{\sf true}}
\newcommand{\false}{\mathrel{\sf false}}

\makeindex
\usepackage{trace}

\documentset{Tokeneer ID Station}

\reference{S.P1229.40.4}

\title{Security Properties}

\synopsis{
This document presents a formal specification of the Security
Properties of the TIS core and a demonstation of satisfaction of these
security properties by the specified TIS core. 
}

\contentsinside

\status{Definitive}

\issuenumber{1.1}

\date{14th August 2008}

\copiedto{%
\begin{minipage}[t]{100mm}%
\vspace{-\abovedisplayskip}%
\begin{multicols}{2}%
\raggedright%
{\bfseries NSA}\\
Randolph Johnson \\[1ex]
{\bfseries SPRE Inc.}\\
\newpage
{\bfseries Praxis High Integrity Systems}\\
Project Team \\[4ex]
~\\
\end{multicols}\end{minipage}
}
\frontsheet{%
Quality
}

\originators{%
David Cooper}

\approverA{Approver}{Janet Barnes}

\confidentiality{}

\begin{document}

\maketitle
%\include{renames}

%=========================================================================

\documentcontrol
Copyright \copyright (2003) United States Government, as represented by the
Director, National Security Agency. All rights reserved.
This material was originally developed by Praxis High Integrity
Systems Ltd. under contract to the National Security Agency.
\section*{Changes History}
\textit{All issues of this document have been type-checked with \fuzz~and have
  given no errors.}
\begin{description}
\item [Issue 0.1] 
(10th July 2003) First draft issued for comments within Praxis.
\item [Issue 1.0] 
(23rd July 2003) Updated following internal review.
First issue to client for comment.
\item [Issue 1.1] 
(14th August 2008) Updated for public release.
\end{description}

\section*{Changes Forecast}

Updates following client review, if any.

\section*{References}
\begin{thebibliography}{99}
\bibitem{Spivey}
The Z Notation: A Reference Manual, J.M Spivey, Prentice Hall, Second Edition,
1992
\bibitem{SRS}
TIS Software Requirements Specification, Version 2.0, S.P1229.41.1.
\bibitem{PP}
TIS Kernel Protection Profile, SPRE Inc, Version 1.0, 5 February 2003.
\bibitem{ST}
TIS Security Target, S.P1229.40.1.
\bibitem{FS}
TIS Formal Specification, S.P1229.50.1.
\end{thebibliography}

\section*{Abbreviations}
{\footnotesize\begin{tabular}{ll}
TIS          & Token ID Station \\
\end{tabular}}

%==========================================================================


\tableofcontents


\include{40_4_SecurityProperties}

%===========================================================================


\end{document}
