\documentclass{pxcsdoc}
\usepackage{praxfuzz}
\usepackage{parskip}
\usepackage{multicol}
\usepackage{enumerate}
\usepackage{proof}


\newcommand{\true}{\mathrel{\sf true}}
\newcommand{\false}{\mathrel{\sf false}}

\makeindex
\usepackage{trace}

\documentset{Tokeneer ID Station}

\reference{S.P1229.41.2}

\title{Formal Specification}

\synopsis{
This document is the formal specification of the core
Token ID Station (TIS), which forms part of Tokeneer. 
}

\contentsinside

\status{Definitive}

\issuenumber{1.4}

\date{14th August 2008}

\copiedto{%
\begin{minipage}[t]{100mm}%
\vspace{-\abovedisplayskip}%
\begin{multicols}{2}%
\raggedright%
{\bfseries NSA}\\
Randolph Johnson \\[1ex]
{\bfseries SPRE Inc.}\\
\newpage
{\bfseries Praxis High Integrity Systems}\\
Project Team \\[4ex]
~\\
\end{multicols}\end{minipage}
}
\frontsheet{%
Quality
}

\originators{%
Janet Barnes}

\approverA{Approver}{David Cooper}

\confidentiality{}

\begin{document}

\maketitle
%\include{renames}

%=========================================================================

\documentcontrol
Copyright \copyright (2003) United States Government, as represented by the
Director, National Security Agency. All rights reserved.
This material was originally developed by Praxis High Integrity
Systems Ltd. under contract to the National Security Agency.
\section*{Changes History}
\textit{All issues of this document have been type-checked with \fuzz~and have
  given no errors.}
\begin{description}
\item [Issue 0.1] 
(28th March 2003) First draft issued for comments within Praxis.
\item [Issue 0.2]
(5th April 2003) Draft incorporating comments from David Cooper.
\item [Issue 0.3]
(14th April 2003) Draft sent to Randolph Johnson (NSA).
\item [Issue 0.4]
(8th May 2003) Final Draft for formal review.
\item [Issue 1.0]
(9th May 2003) Provisional issue to client following internal review.
\item [Issue 1.1]
(27th June 2003) Definitive issue incorporating client feedback and
correction to faults S.P1229.6.1-5. 
\item [Issue 1.2]
(22nd July 2003) Definitive issue correcting faults 
\begin{itemize}
\item S.P1229.6.6 - Tearing a user token mid-entry should constitute a
failed entry and be logged as such within the statistics.
\item S.P1229.6.7 - Remove the ability to retry fingerprint validation
following a failure.
\item S.P1229.6.19 - Correct typographical error on page 51.
\end{itemize}
\item [Issue 1.3]
(22nd August 2003) Definitive issue correcting faults.
\begin{itemize}
\item S.P1229.6.32 - Improve poor text messages on screen.
\item S.P1229.6.33 - Make initial configuration realistic.
\item S.P1229.6.36 - Screen should show busy message when a user entry
is in progress.
\item S.P1229.6.38 - Operation failures not reported on screen.
\end{itemize}
\item [Issue 1.4]
(14th August 2008) Updated for public release.
\end{description}

\section*{Changes Forecast}

None. This document is now under change control.

\section*{References}
\begin{thebibliography}{99}
\bibitem{Spivey}
The Z Notation: A Reference Manual, J.M Spivey, Prentice Hall, Second Edition,
1992
\bibitem{SRS}
TIS Software Requirements Specification, Version 2.0, S.P1229.41.1.
\bibitem{PP}
TIS Kernel Protection Profile, SPRE Inc, Version 1.0, 5 February 2003.
\bibitem{ST}
TIS Security Target, S.P1229.40.1.
\end{thebibliography}

\section*{Abbreviations}
{\footnotesize\begin{tabular}{ll}
AA           & Attribute Authority \\
ATR          & Answer-to-Reset \\
CA           & Certification Authority \\
I\&A         & Identification and Authentication \\
RSA          & Rivest Shamir Adelman algorithm \\   
SPARK        & SPADE Ada Kernel (analysable Ada subset from Praxis) \\
SRS          & Software Requirements Specification \\
TIS          & Token ID Station \\
\end{tabular}}

%==========================================================================


\tableofcontents


\include{41_2_Intro}

\include{41_2_TIS}

\include{41_2_Interfaces}

\include{41_2_Internal}

\include{41_2_UserEntry}

\include{41_2_Enclave}

\include{41_2_Start}

\include{41_2_Whole}

\appendix

\input{41_2_Zbasics}

\input{41_2_Summary}

\input{41_2_Pre}

\input{41_2_SRSTrace}

\input{41_2_STTrace}

\chapter{Z index}

This section contains an index of Z terms.
This contains all the Z schemas, types and functions
defined in the specification. 

\begin{multicols}{2}\scriptsize
\begin{theindex}
\input{z-index}
\end{theindex}
\end{multicols}

\chapter{Traceunit index}

An index of traceunits. This contains all the traceunits placed in the
specification to enable the elements of the specification to be traced to
the design.

\begin{multicols}{2}\scriptsize
\begin{theindex}
\input{trace-index}
\end{theindex}
\end{multicols}

\chapter{Requirements index}

An index of traceunits. This contains all the traceunits in the 
requirements documents . All requirements are listed with the pages
from which they are referenced. 

\begin{multicols}{2}\scriptsize
\begin{theindex}
\input{req-index}
\end{theindex}
\end{multicols}

%===========================================================================


\end{document}
